### 2021-03-28 control.tex
working on control

% Created by tikzDevice version 0.12.3 on 2021-03-27 20:28:52
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (722.70,361.35);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,361.35);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.7pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (361.35,361.35);
\end{scope}
\begin{scope}
\path[clip] ( 49.41, 39.63) rectangle (354.35,332.06);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 49.41, 39.63) rectangle (354.35,332.06);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 63.27, 52.92) --
	( 72.51, 59.52) --
	( 81.75, 66.11) --
	( 90.99, 72.69) --
	(100.23, 79.24) --
	(109.48, 85.78) --
	(118.72, 92.31) --
	(127.96, 98.81) --
	(137.20,105.30) --
	(146.44,111.78) --
	(155.68,118.24) --
	(164.92,124.68) --
	(174.16,131.11) --
	(183.40,137.52) --
	(192.64,143.92) --
	(201.88,150.30) --
	(211.12,156.67) --
	(220.36,163.02) --
	(229.60,169.35) --
	(238.84,175.67) --
	(248.08,181.98) --
	(257.32,188.27) --
	(266.56,194.54) --
	(275.81,200.80) --
	(285.05,207.04) --
	(294.29,213.27) --
	(303.53,219.49) --
	(312.77,225.69) --
	(322.01,231.88) --
	(331.25,238.05) --
	(340.49,244.20);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 63.27,205.70) --
	( 72.51,205.70) --
	( 81.75,205.70) --
	( 90.99,205.70) --
	(100.23,205.70) --
	(109.48,205.70) --
	(118.72,205.70) --
	(127.96,205.70) --
	(137.20,205.70) --
	(146.44,205.70) --
	(155.68,205.70) --
	(164.92,205.70) --
	(174.16,205.70) --
	(183.40,205.70) --
	(192.64,205.70) --
	(201.88,205.70) --
	(211.12,205.70) --
	(220.36,205.70) --
	(229.60,205.70) --
	(238.84,205.70) --
	(248.08,205.70) --
	(257.32,205.70) --
	(266.56,205.70) --
	(275.81,205.70) --
	(285.05,205.70) --
	(294.29,205.70) --
	(303.53,205.70) --
	(312.77,205.70) --
	(322.01,205.70) --
	(331.25,205.70) --
	(340.49,205.70);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt ,line join=round] ( 63.27, 87.55) --
	( 72.51, 95.59) --
	( 81.75,103.61) --
	( 90.99,111.61) --
	(100.23,119.58) --
	(109.48,127.53) --
	(118.72,135.45) --
	(127.96,143.35) --
	(137.20,151.22) --
	(146.44,159.07) --
	(155.68,166.90) --
	(164.92,174.71) --
	(174.16,182.49) --
	(183.40,190.25) --
	(192.64,197.99) --
	(201.88,205.70) --
	(211.12,213.39) --
	(220.36,221.06) --
	(229.60,228.71) --
	(238.84,236.33) --
	(248.08,243.93) --
	(257.32,251.51) --
	(266.56,259.07) --
	(275.81,266.61) --
	(285.05,274.12) --
	(294.29,281.62) --
	(303.53,289.09) --
	(312.77,296.54) --
	(322.01,303.97) --
	(331.25,311.38) --
	(340.49,318.77);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.7pt,line join=round,line cap=round] ( 49.41, 39.63) rectangle (354.35,332.06);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.12] at ( 43.11,109.07) {1.50};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.12] at ( 43.11,190.12) {1.75};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.12] at ( 43.11,260.33) {2.00};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.12] at ( 43.11,322.26) {2.25};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.7pt,line join=round] ( 45.91,112.92) --
	( 49.41,112.92);

\path[draw=drawColor,line width= 0.7pt,line join=round] ( 45.91,193.98) --
	( 49.41,193.98);

\path[draw=drawColor,line width= 0.7pt,line join=round] ( 45.91,264.19) --
	( 49.41,264.19);

\path[draw=drawColor,line width= 0.7pt,line join=round] ( 45.91,326.12) --
	( 49.41,326.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.7pt,line join=round] ( 63.27, 36.13) --
	( 63.27, 39.63);

\path[draw=drawColor,line width= 0.7pt,line join=round] (132.58, 36.13) --
	(132.58, 39.63);

\path[draw=drawColor,line width= 0.7pt,line join=round] (201.88, 36.13) --
	(201.88, 39.63);

\path[draw=drawColor,line width= 0.7pt,line join=round] (271.18, 36.13) --
	(271.18, 39.63);

\path[draw=drawColor,line width= 0.7pt,line join=round] (340.49, 36.13) --
	(340.49, 39.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.12] at ( 63.27, 25.61) {0.6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.12] at (132.58, 25.61) {0.8};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.12] at (201.88, 25.61) {1.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.12] at (271.18, 25.61) {1.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.12] at (340.49, 25.61) {1.4};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.40] at (201.88, 10.06) {Generation interval ratio, $\bar{G}_{\mathrm{var}}/\bar{G}_{\mathrm{wt}}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.40] at ( 16.64,185.84) {Strength};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 54.94,251.68) rectangle (226.84,324.71);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 61.94,293.37) rectangle ( 79.29,310.71);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt ,line join=round] ( 63.68,302.04) -- ( 77.55,302.04);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt ,line join=round] ( 63.68,302.04) -- ( 77.55,302.04);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt ,line join=round] ( 63.68,302.04) -- ( 77.55,302.04);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 61.94,276.02) rectangle ( 79.29,293.37);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 63.68,284.69) -- ( 77.55,284.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 63.68,284.69) -- ( 77.55,284.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 63.68,284.69) -- ( 77.55,284.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 61.94,258.68) rectangle ( 79.29,276.02);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 63.68,267.35) -- ( 77.55,267.35);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 63.68,267.35) -- ( 77.55,267.35);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 63.68,267.35) -- ( 77.55,267.35);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.12] at ( 86.29,298.18) {Constant speed intervention};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.12] at ( 86.29,280.84) {Constant strength intervention};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.12] at ( 86.29,263.49) {Epidemic};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.68] at ( 49.41,342.78) {A};
\end{scope}
\begin{scope}
\path[clip] (361.35,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.7pt,line join=round,line cap=round,fill=fillColor] (361.35,  0.00) rectangle (722.70,361.35);
\end{scope}
\begin{scope}
\path[clip] (416.36, 39.63) rectangle (715.70,332.06);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (416.36, 39.63) rectangle (715.70,332.06);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (429.97, 79.69) --
	(439.04, 79.69) --
	(448.11, 79.69) --
	(457.18, 79.69) --
	(466.25, 79.69) --
	(475.32, 79.69) --
	(484.39, 79.69) --
	(493.46, 79.69) --
	(502.53, 79.69) --
	(511.60, 79.69) --
	(520.68, 79.69) --
	(529.75, 79.69) --
	(538.82, 79.69) --
	(547.89, 79.69) --
	(556.96, 79.69) --
	(566.03, 79.69) --
	(575.10, 79.69) --
	(584.17, 79.69) --
	(593.24, 79.69) --
	(602.31, 79.69) --
	(611.38, 79.69) --
	(620.46, 79.69) --
	(629.53, 79.69) --
	(638.60, 79.69) --
	(647.67, 79.69) --
	(656.74, 79.69) --
	(665.81, 79.69) --
	(674.88, 79.69) --
	(683.95, 79.69) --
	(693.02, 79.69) --
	(702.09, 79.69);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] (429.97,318.77) --
	(439.04,298.97) --
	(448.11,280.79) --
	(457.18,264.03) --
	(466.25,248.54) --
	(475.32,234.18) --
	(484.39,220.82) --
	(493.46,208.37) --
	(502.53,196.74) --
	(511.60,185.84) --
	(520.68,175.62) --
	(529.75,166.00) --
	(538.82,156.95) --
	(547.89,148.40) --
	(556.96,140.32) --
	(566.03,132.67) --
	(575.10,125.42) --
	(584.17,118.54) --
	(593.24,112.00) --
	(602.31,105.77) --
	(611.38, 99.83) --
	(620.46, 94.17) --
	(629.53, 88.76) --
	(638.60, 83.59) --
	(647.67, 78.65) --
	(656.74, 73.91) --
	(665.81, 69.36) --
	(674.88, 65.00) --
	(683.95, 60.82) --
	(693.02, 56.79) --
	(702.09, 52.92);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt ,line join=round] (429.97,139.21) --
	(439.04,139.21) --
	(448.11,139.21) --
	(457.18,139.21) --
	(466.25,139.21) --
	(475.32,139.21) --
	(484.39,139.21) --
	(493.46,139.21) --
	(502.53,139.21) --
	(511.60,139.21) --
	(520.68,139.21) --
	(529.75,139.21) --
	(538.82,139.21) --
	(547.89,139.21) --
	(556.96,139.21) --
	(566.03,139.21) --
	(575.10,139.21) --
	(584.17,139.21) --
	(593.24,139.21) --
	(602.31,139.21) --
	(611.38,139.21) --
	(620.46,139.21) --
	(629.53,139.21) --
	(638.60,139.21) --
	(647.67,139.21) --
	(656.74,139.21) --
	(665.81,139.21) --
	(674.88,139.21) --
	(683.95,139.21) --
	(693.02,139.21) --
	(702.09,139.21);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.7pt,line join=round,line cap=round] (416.36, 39.63) rectangle (715.70,332.06);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.12] at (410.06, 75.84) {0.100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.12] at (410.06,139.36) {0.125};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.12] at (410.06,202.88) {0.150};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.12] at (410.06,266.39) {0.175};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.7pt,line join=round] (412.86, 79.69) --
	(416.36, 79.69);

\path[draw=drawColor,line width= 0.7pt,line join=round] (412.86,143.21) --
	(416.36,143.21);

\path[draw=drawColor,line width= 0.7pt,line join=round] (412.86,206.73) --
	(416.36,206.73);

\path[draw=drawColor,line width= 0.7pt,line join=round] (412.86,270.25) --
	(416.36,270.25);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.7pt,line join=round] (429.97, 36.13) --
	(429.97, 39.63);

\path[draw=drawColor,line width= 0.7pt,line join=round] (498.00, 36.13) --
	(498.00, 39.63);

\path[draw=drawColor,line width= 0.7pt,line join=round] (566.03, 36.13) --
	(566.03, 39.63);

\path[draw=drawColor,line width= 0.7pt,line join=round] (634.06, 36.13) --
	(634.06, 39.63);

\path[draw=drawColor,line width= 0.7pt,line join=round] (702.09, 36.13) --
	(702.09, 39.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.12] at (429.97, 25.61) {0.6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.12] at (498.00, 25.61) {0.8};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.12] at (566.03, 25.61) {1.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.12] at (634.06, 25.61) {1.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.12] at (702.09, 25.61) {1.4};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.40] at (566.03, 10.06) {Generation interval ratio, $\bar{G}_{\mathrm{var}}/\bar{G}_{\mathrm{wt}}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.40] at (377.99,185.84) {Speed (1/days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.68] at (416.36,342.78) {B};
\end{scope}
\end{tikzpicture}

\end{document}