### 2021-04-02 control.tex

% Created by tikzDevice version 0.12.3 on 2021-03-28 18:34:28
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (722.70,361.35);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.12] at ( 37.52, 56.08) {1.4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.12] at ( 37.52,123.48) {1.6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.12] at ( 37.52,182.94) {1.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.12] at ( 37.52,236.12) {2.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.12] at ( 37.52,284.23) {2.2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.12] at ( 37.52,328.15) {2.4};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.7pt,line join=round] ( 40.32, 59.94) --
	( 43.82, 59.94);

\path[draw=drawColor,line width= 0.7pt,line join=round] ( 40.32,127.34) --
	( 43.82,127.34);

\path[draw=drawColor,line width= 0.7pt,line join=round] ( 40.32,186.79) --
	( 43.82,186.79);

\path[draw=drawColor,line width= 0.7pt,line join=round] ( 40.32,239.98) --
	( 43.82,239.98);

\path[draw=drawColor,line width= 0.7pt,line join=round] ( 40.32,288.09) --
	( 43.82,288.09);

\path[draw=drawColor,line width= 0.7pt,line join=round] ( 40.32,332.01) --
	( 43.82,332.01);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.40] at ( 16.64,185.89) {Strength};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.12] at (404.47, 94.91) {0.100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.12] at (404.47,167.50) {0.125};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.12] at (404.47,240.08) {0.150};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.12] at (404.47,312.67) {0.175};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.7pt,line join=round] (407.27, 98.77) --
	(410.77, 98.77);

\path[draw=drawColor,line width= 0.7pt,line join=round] (407.27,171.35) --
	(410.77,171.35);

\path[draw=drawColor,line width= 0.7pt,line join=round] (407.27,243.94) --
	(410.77,243.94);

\path[draw=drawColor,line width= 0.7pt,line join=round] (407.27,316.52) --
	(410.77,316.52);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.40] at (372.39,185.89) {Speed (1/days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.68] at ( 43.82,342.78) {A};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.68] at (410.77,342.78) {B};
\end{scope}
\begin{scope}
\path[clip] ( 43.82, 39.69) rectangle (348.75,332.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 43.82, 39.69) rectangle (348.75,332.09);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 57.68, 52.98) --
	( 66.92, 57.30) --
	( 76.16, 61.73) --
	( 85.40, 66.27) --
	( 94.64, 70.93) --
	(103.89, 75.71) --
	(113.13, 80.60) --
	(122.37, 85.63) --
	(131.61, 90.78) --
	(140.85, 96.06) --
	(150.09,101.47) --
	(159.33,107.02) --
	(168.57,112.71) --
	(177.81,118.54) --
	(187.05,124.52) --
	(196.29,130.64) --
	(205.53,136.92) --
	(214.77,143.36) --
	(224.01,149.95) --
	(233.25,156.71) --
	(242.49,163.63) --
	(251.73,170.73) --
	(260.97,177.99) --
	(270.21,185.44) --
	(279.45,193.06) --
	(288.69,200.87) --
	(297.93,208.87) --
	(307.17,217.06) --
	(316.41,225.45) --
	(325.65,234.04) --
	(334.89,242.84);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 57.68,183.83) --
	( 66.92,183.83) --
	( 76.16,183.83) --
	( 85.40,183.83) --
	( 94.64,183.83) --
	(103.89,183.83) --
	(113.13,183.83) --
	(122.37,183.83) --
	(131.61,183.83) --
	(140.85,183.83) --
	(150.09,183.83) --
	(159.33,183.83) --
	(168.57,183.83) --
	(177.81,183.83) --
	(187.05,183.83) --
	(196.29,183.83) --
	(205.53,183.83) --
	(214.77,183.83) --
	(224.01,183.83) --
	(233.25,183.83) --
	(242.49,183.83) --
	(251.73,183.83) --
	(260.97,183.83) --
	(270.21,183.83) --
	(279.45,183.83) --
	(288.69,183.83) --
	(297.93,183.83) --
	(307.17,183.83) --
	(316.41,183.83) --
	(325.65,183.83) --
	(334.89,183.83);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt ,line join=round] ( 57.68, 89.66) --
	( 66.92, 94.92) --
	( 76.16,100.30) --
	( 85.40,105.82) --
	( 94.64,111.48) --
	(103.89,117.28) --
	(113.13,123.23) --
	(122.37,129.32) --
	(131.61,135.57) --
	(140.85,141.97) --
	(150.09,148.53) --
	(159.33,155.25) --
	(168.57,162.14) --
	(177.81,169.19) --
	(187.05,176.42) --
	(196.29,183.83) --
	(205.53,191.42) --
	(214.77,199.19) --
	(224.01,207.14) --
	(233.25,215.29) --
	(242.49,223.64) --
	(251.73,232.19) --
	(260.97,240.94) --
	(270.21,249.90) --
	(279.45,259.07) --
	(288.69,268.46) --
	(297.93,278.07) --
	(307.17,287.90) --
	(316.41,297.96) --
	(325.65,308.26) --
	(334.89,318.80);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.7pt,line join=round,line cap=round] ( 43.82, 39.69) rectangle (348.75,332.09);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 48.06,251.71) rectangle (222.54,324.75);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 55.06,293.40) rectangle ( 72.41,310.75);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt ,line join=round] ( 56.80,302.08) -- ( 70.67,302.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt ,line join=round] ( 56.80,302.08) -- ( 70.67,302.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt ,line join=round] ( 56.80,302.08) -- ( 70.67,302.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 55.06,276.06) rectangle ( 72.41,293.40);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 56.80,284.73) -- ( 70.67,284.73);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 56.80,284.73) -- ( 70.67,284.73);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 56.80,284.73) -- ( 70.67,284.73);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 55.06,258.71) rectangle ( 72.41,276.06);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 56.80,267.39) -- ( 70.67,267.39);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 56.80,267.39) -- ( 70.67,267.39);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 56.80,267.39) -- ( 70.67,267.39);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.12] at ( 79.41,298.22) {Constant speed intervention};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.12] at ( 79.41,280.87) {Constant strength intervention};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.12] at ( 79.41,263.53) {Epidemic};
\end{scope}
\begin{scope}
\path[clip] (410.77, 39.69) rectangle (715.70,332.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (410.77, 39.69) rectangle (715.70,332.09);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (424.63, 98.77) --
	(433.87, 98.77) --
	(443.11, 98.77) --
	(452.35, 98.77) --
	(461.59, 98.77) --
	(470.83, 98.77) --
	(480.07, 98.77) --
	(489.31, 98.77) --
	(498.55, 98.77) --
	(507.79, 98.77) --
	(517.03, 98.77) --
	(526.28, 98.77) --
	(535.52, 98.77) --
	(544.76, 98.77) --
	(554.00, 98.77) --
	(563.24, 98.77) --
	(572.48, 98.77) --
	(581.72, 98.77) --
	(590.96, 98.77) --
	(600.20, 98.77) --
	(609.44, 98.77) --
	(618.68, 98.77) --
	(627.92, 98.77) --
	(637.16, 98.77) --
	(646.40, 98.77) --
	(655.64, 98.77) --
	(664.88, 98.77) --
	(674.12, 98.77) --
	(683.36, 98.77) --
	(692.60, 98.77) --
	(701.84, 98.77);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] (424.63,318.80) --
	(433.87,306.04) --
	(443.11,293.62) --
	(452.35,281.53) --
	(461.59,269.76) --
	(470.83,258.31) --
	(480.07,247.16) --
	(489.31,236.31) --
	(498.55,225.75) --
	(507.79,215.47) --
	(517.03,205.47) --
	(526.28,195.73) --
	(535.52,186.25) --
	(544.76,177.03) --
	(554.00,168.05) --
	(563.24,159.31) --
	(572.48,150.80) --
	(581.72,142.52) --
	(590.96,134.46) --
	(600.20,126.62) --
	(609.44,118.98) --
	(618.68,111.55) --
	(627.92,104.32) --
	(637.16, 97.28) --
	(646.40, 90.42) --
	(655.64, 83.75) --
	(664.88, 77.26) --
	(674.12, 70.94) --
	(683.36, 64.79) --
	(692.60, 58.81) --
	(701.84, 52.98);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt ,line join=round] (424.63,166.78) --
	(433.87,166.78) --
	(443.11,166.78) --
	(452.35,166.78) --
	(461.59,166.78) --
	(470.83,166.78) --
	(480.07,166.78) --
	(489.31,166.78) --
	(498.55,166.78) --
	(507.79,166.78) --
	(517.03,166.78) --
	(526.28,166.78) --
	(535.52,166.78) --
	(544.76,166.78) --
	(554.00,166.78) --
	(563.24,166.78) --
	(572.48,166.78) --
	(581.72,166.78) --
	(590.96,166.78) --
	(600.20,166.78) --
	(609.44,166.78) --
	(618.68,166.78) --
	(627.92,166.78) --
	(637.16,166.78) --
	(646.40,166.78) --
	(655.64,166.78) --
	(664.88,166.78) --
	(674.12,166.78) --
	(683.36,166.78) --
	(692.60,166.78) --
	(701.84,166.78);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.7pt,line join=round,line cap=round] (410.77, 39.69) rectangle (715.70,332.09);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.7pt,line join=round] ( 59.39, 36.19) --
	( 59.39, 39.69);

\path[draw=drawColor,line width= 0.7pt,line join=round] (196.29, 36.19) --
	(196.29, 39.69);

\path[draw=drawColor,line width= 0.7pt,line join=round] (334.89, 36.19) --
	(334.89, 39.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.12] at ( 59.39, 25.68) {0.67};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.12] at (196.29, 25.68) {1.00};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.12] at (334.89, 25.68) {1.50};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.40] at (196.29, 10.07) {Generation interval ratio, $\bar{G}_{\mathrm{var}}/\bar{G}_{\mathrm{wt}}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.7pt,line join=round] (426.34, 36.19) --
	(426.34, 39.69);

\path[draw=drawColor,line width= 0.7pt,line join=round] (563.24, 36.19) --
	(563.24, 39.69);

\path[draw=drawColor,line width= 0.7pt,line join=round] (701.84, 36.19) --
	(701.84, 39.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.12] at (426.34, 25.68) {0.67};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.12] at (563.24, 25.68) {1.00};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.12] at (701.84, 25.68) {1.50};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.40] at (563.24, 10.07) {Generation interval ratio, $\bar{G}_{\mathrm{var}}/\bar{G}_{\mathrm{wt}}$};
\end{scope}
\end{tikzpicture}

\end{document}